Investigating Cross-Sectional Size Distribution of Randomly Distributed 3D Spheres

### 2020-05-31 Paper/main.tex
Minor changes

\IEEEPARstart{I}{n} the world of Material Science and Crystallography, it is
often the case that a material is identified and classified using the
distribution of its constituent particle sizes. However, direct observation of
the particles so as to construct a distribution is often impossible. Using a
modern microscope or laser imaging techniques, it is possible to observe
particles in a cross-section of material. But the observed radii of particles in
this cross-section do not necessarily correspond directly to the radii of the
particles in the material, since the plane of the cross-section is not
necessarily oriented such that the observed particles lie exactly on the
plane. Thus, it becomes difficult to relate the observed distribution of
particle sizes in the plane to the real distribution of particle sizes in the
material.

Mathematically, determining the distribution of spherical particles from planar
sections is a well known problem called Wicksell's Corpuscle Problem
\cite{DWicksell}, named after S. D. Wicksell, who solved the problem in 1925. With
this solution, it is possible to determine the original distribution of
spherical particles. Thus, it is possible to determine the distribution of
particle sizes in a solid from its observed cross-section, allowing us to
identify and classify materials better \cite{Cuzzi2016}.

In this paper we aim to show that it is possible to reproduce theoretical
results on cross-sectional distributions using numerical simulation, and it is
thus possible to solve Wicksell's Corpuscle Problem in this manner.

@@

Consider many spheres of different radii distributed randomly in space. For the
purposes of the numerical simulation, we will consider a cube of length 20. The
spheres were not allowed to overlap, which we ensured by placing the spheres such
that if one sphere has vector position $\overrightarrow{\textbf{a}}$ and radius
$r_{1}$ and the other has vector position $\overrightarrow{\textbf{b}}$ and
radius $r_{2}$, the spheres must have positions such that:
\begin{equation*}
|\overrightarrow{\textbf{a}} - \overrightarrow{\textbf{b}}| > (r_{1} + r_{2})
\end{equation*}
Random coordinates for a sphere are drawn from a uniform distribution and
checked against every existing sphere until a set of coordinates is drawn that
ensures no overlap, at which point the sphere is placed in space and the process
is repeated until all the necessary spheres have been placed.

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/box3d.pdf}
\caption{$750$ spheres of radius 0.7 distributed randomly in a cube of length 20,
  such that no sphere is touching or overlapping another sphere. A randomly
  oriented plane passes through the box and ``cuts'' through the
  spheres.}\label{fig:3dplot_plane}
\end{center}
\end{figure}

We then placed a plane into the cube, with a random orientation. The plane will
``cut'' through many of the spheres in the cube. This is shown in
Figure~\ref{fig:3dplot_plane}. Here we see $750$ spheres, each with radius 0.7
distributed randomly in a cube of length 20. The plane is shown in blue, and can
be seen intersecting many of the spheres in the cube.

In order to obtain the planar cross-section, we employed the spherical cap
method \cite{Sykes1922}. If we consider a sphere with centre
$\overrightarrow{\textbf{c}_0} = (x_0, y_0, z_0)$ and radius $R$, and a plane
with equation $Ax + By + Cz = D$ such that $\overrightarrow{\textbf{n}} = (A, B,
C)$ is the normal vector to the plane, then the distance from the centre of the
circle to the closest point on the plane $p_0$ is given by $\rho$:
\begin{equation*}
\rho = \frac{(\overrightarrow{\textbf{c}_0} - \overrightarrow{\textbf{p}_0}) \cdot{}
  \overrightarrow{\textbf{n}}}{|\overrightarrow{\textbf{n}}|} = \frac{Ax_0 +
  By_0 + Cz_0 - D}{\sqrt{A^2 + B^2 + C^2}}
\end{equation*}
If $|\rho| < R$, then there is an intersection of the plane and the sphere. The
radius of the circle $r$ in the planar cross-section is given by:
\begin{equation*}
r = \sqrt{R^2 - \rho^2}
\end{equation*}
And we then binned this radius to produce a histogram of the distribution of
planar circle radii. It is also possible to recover the centre
$\overrightarrow{\textbf{c}}$ of the circle visible on the plane using:
\begin{equation*}
\overrightarrow{\textbf{c}} = \overrightarrow{\textbf{c}_0} +
\rho\frac{\overrightarrow{\textbf{n}}}{|\overrightarrow{\textbf{c}}|} = (x_0,
y_0, z_0) + \rho\frac{(A, B, C)}{\sqrt{A^2 + B^2 + C^2}}
\end{equation*}

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/circles.pdf}
\caption{Planar cross-section constructed using mock data. Spheres cut by the
  plane appear as circles on the plane, with radius dependent on the relative
  position of the spheres to the plane. This distribution of circles can be
  modelled.}\label{fig:circles}
\end{center}
\end{figure}

@@

If we compute the integral in Equation~\ref{eq:phi_OG}, replacing $f(R)$ with
$\delta(R-R_0)$ we find:
\begin{equation}
\phi(x) = \int_{x}^{\infty}\frac{x\delta(R-R_0)}{\sqrt{R^2 - x^2}} =
\frac{xH(R_0-x)}{\sqrt{R_0^2-x^2}}
\label{eq:phi_constant}
\end{equation}
where $H(R_0 - x)$ is the Heaviside function, which ``turns on'' the
distribution only for $x < R0$ as it is not possible to have a circular radius
$x$ larger than the spherical radius. So we aim to see our experimental data
reproduce this distribution \cite{Kong2005}. We also investigated how changing the
value of $R_0$ affects the experimental distribution.

@@

We next randomised the radius of each sphere to be in the range $0.65 < R <
0.75$. The experimental data for this simulation is presented in
Figure~\ref{fig:random}.

>>>>

\IEEEPARstart{I}{n} the world of Material Science and Crystallography, it is
often the case that a material is identified and classified using the
distribution of its constituent particle sizes. However, direct observation of
the particles is often impossible. Using a modern microscope or laser imaging
techniques, it is possible to observe particles in a cross-section of
material. But the observed radii of particles in this cross-section do not
necessarily correspond directly to the radii of the particles in the material,
since the plane is not oriented such that it passes through the centre of all
observed particles. Thus, it becomes difficult to relate the observed
distribution of particle sizes in the plane to the real distribution of particle
sizes in the material.

Mathematically, determining the distribution of spherical particles from planar
sections is a well known problem called Wicksell's Corpuscle Problem
\cite{DWicksell}, named after S. D. Wicksell, who solved the problem in
1925. With this solution, it is possible to determine the distribution of
particle sizes in a solid from its observed cross-section, allowing us to
identify and classify materials better \cite{Cuzzi2016}.

In this paper we aim to show that it is possible to reproduce theoretical
results for cross-sectional distributions using numerical simulation, and it is
thus possible to solve Wicksell's Corpuscle Problem in this manner.

@@

The numerical simulation of the 3D particle distribution was done using
Python. It looked at many spheres of different radii, distributed randomly in
space with a given "box of length 20. The spheres were not allowed to overlap,
which we ensured by placing the spheres such that if one sphere has vector
position $\overrightarrow{\textbf{a}}$ and radius $r_{1}$ and the other has
vector position $\overrightarrow{\textbf{b}}$ and radius $r_{2}$, the spheres
must have positions such that:
\begin{equation*}
|\overrightarrow{\textbf{a}} - \overrightarrow{\textbf{b}}| > (r_{1} + r_{2})
\end{equation*}
Random coordinates for a sphere were drawn from a uniform distribution and
checked against every existing sphere until a set of coordinates was drawn that
ensures no overlap, at which point the sphere was placed in space and the process
was repeated until all the necessary spheres have been placed.

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/box3d.pdf}
\caption{$750$ spheres of radius 0.7 distributed randomly in a cube of length 20,
  such that no sphere is touching or overlapping another sphere. A horizontal
  plane, in blue, passes through the box and ``cuts'' through the
  spheres.}\label{fig:3dplot_plane}
\end{center}
\end{figure}

We then placed a plane into the cube, with a random orientation, which
``cuts'' through many of the spheres in the cube. This is shown in
Figure~\ref{fig:3dplot_plane}, where we see a plane (in blue) intersecting 750
spheres of constant radius, distributed at random in a box.

In order to obtain the planar cross-section, we employed the spherical cap
method \cite{Sykes1922}. If we consider a sphere with centre
$\overrightarrow{\textbf{c}_0} = (x_0, y_0, z_0)$ and radius $R$, and a plane
with equation $Ax + By + Cz = D$ such that $\overrightarrow{\textbf{n}} = (A, B,
C)$ is the normal vector to the plane, then the distance from the centre of the
circle to the closest point on the plane $\overrightarrow{\textbf{p}_0}$ is given by $\rho$:
\begin{equation*}
\rho = \frac{(\overrightarrow{\textbf{c}_0} - \overrightarrow{\textbf{p}_0}) \cdot{}
  \overrightarrow{\textbf{n}}}{|\overrightarrow{\textbf{n}}|} = \frac{Ax_0 +
  By_0 + Cz_0 - D}{\sqrt{A^2 + B^2 + C^2}}
\end{equation*}
If $|\rho| < R$, then there is an intersection of the plane and the sphere. The
radius of the circle $r$ in the planar cross-section is given by:
\begin{equation*}
r = \sqrt{R^2 - \rho^2}
\end{equation*}
And we then binned this radius to produce a histogram of the distribution of
planar circle radii. It is also possible to recover the centre
$\overrightarrow{\textbf{c}}$ of the circle visible on the plane using:
\begin{equation*}
\overrightarrow{\textbf{c}} = \overrightarrow{\textbf{c}_0} +
\rho\frac{\overrightarrow{\textbf{n}}}{|\overrightarrow{\textbf{c}}|} = (x_0,
y_0, z_0) + \rho\frac{(A, B, C)}{\sqrt{A^2 + B^2 + C^2}}
\end{equation*}

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/circles.pdf}
\caption{Planar cross-section constructed using mock data. Spheres cut by the
  plane appear as circles on the plane, with radius dependent on the relative
  position of the spheres to the plane. This distribution of the radii of the
  circles can be modelled.}\label{fig:circles}
\end{center}
\end{figure}

@@

If we compute the integral in Equation~\ref{eq:phi_OG}, replacing $f(R)$ with
$\delta(R-R_0)$ we find:
\begin{equation}
\phi(x) = \int_{x}^{\infty}\frac{x\delta(R-R_0)}{\sqrt{R^2 - x^2}} =
\frac{xH(R_0-x)}{\sqrt{R_0^2-x^2}}
\label{eq:phi_constant}
\end{equation}
where $H(R_0 - x)$ is the Heaviside function, which ``turns on'' the
distribution only for $x < R_0$ as it is not possible to have a circular radius
$x$ larger than the spherical radius. So we aim to see our experimental data
reproduce this distribution \cite{Kong2005}. We also investigated how changing
the size of the spheres used in the simulation affects the experimental data.

@@
